Discriminative Sensing Based on Signal Processing for Information Security Analysis

### 2018-05-15 docs/thesis/chapters/1_introduction/introduction.tex
new project for adaptive multivariage gaussian applied to network attack detection

The field of Mobile Money Transfer (MMT) is a growing market segment and has been motivating investiments into security defense and fraud detection. Obtaining access to data sets of mobile transactions for research is a very hard task due to the intrinsic private nature of such transactions. Scientists and researchers must today spend time and effort in obtaining access to relevant data sets before they can research on such data set. However, PaySim \cite{lopez2016paysim} is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.

@@

Considering that imbalanced data can be composed by rare or unknown events, adaptive techniques can be useful in order to improve models and be able to react to changes or unknown patterns and maintain acceptable performance for classification algorithms. Adaptive techniques are also important for online data analysis or streaming processing, where an incremental processing approach can be applied to deal with high thoughput data analysis.

Compressed Sensing (CS) aims to obtain the most relevant information of a dataset, what makes it useful for compression, signal reconstruction, image processing and classification, such as in sparce representation classification (SRC) problems. Dictionary learning is related to CS but seeks to obtain compact and meaningful signal representations, known as dictionary, from learning algorithms or models. DL is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries. DL is based on the principle that some observations can be described by a sparse subset of atoms taken from a redundant dictionary, which represents the causes of some observations of the world. The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, image processing, data decomposition, feature extraction and classification \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.

Dictionaries are either available analytically, or can be learned from a suitable training set. DL can be formulated as an adptive algorithm that learns from traning data that is updated to represent changes or to adpat to new patterns. Iterative dictionary learning has been used as a blind technique for feature extraction, improving classification without require feature selection or principal conponent analysis

@@

Considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, we also propose a tensor-based dictionary learning method for fraud detecton from imbalanced data, in order to apply the tensor deconposition for dictionary learning methods to evidentiate the discriminative sensing of a fraud detection dataset. Specifically, we propose to apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification of mobile money transactions. We propose to use tensor-based dictionary learning for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.

@@

We analyse problems related to detection of information security issues and propose new approaches to improve malicious behavior detection through signal processing techniques. The results of the work presented in this thesis provide the following contributions:

\begin{enumerate}
	\item We propose an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 dataset;
	\item We discuss the computational complexity of the proposed framework and evaluate the required processing time for tested scenarios;
	\item We propose an architechture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
	\item We discuss the processing time of the proposed framework for mobile devices;
	\item We propose a tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
\end{enumerate}

@@

This thesis is organized as follows. In Chapter \ref{ch:2_mos_eig_sim}, we propose a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks, based on eigenvalue analysis, model order selection (MOS) and similarity analysis. Chapter \ref{ch:3_mobile} presents an evaluation of an approach and architecture based on user behavior analysis through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats in a mobile application. Chapter \ref{ch:4_tensor_dl} proposes a tensor-based dictionary learning approach for fraud detection from mobile payment transactions. Chapter \ref{ch:5_conclusionfuturework} draws the conclusions and the suggestions for future work. Futhermore, the Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes and the Appendix \ref{apx:b_csf_fs} presents a critical factors analysis based on Principal Compoment Analytis (PCA) for visual discriminant analysis, and presents an approach based on Recursive Feature Elimination (RFE) combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the Critical Success Factors (CSF) for IT governance of the public sector according to TCU.

>>>>

The field of Mobile Money Transfer (MMT) is a growing market segment and has been motivating investments into security defense and fraud detection. Obtaining access to data sets of mobile transactions for research is a very hard task due to the intrinsic private nature of such transactions. Scientists and researchers must today spend time and effort in obtaining access to relevant data sets before they can research on such data set. However, PaySim \cite{lopez2016paysim} is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.

@@

Considering that imbalanced data can be composed by rare or unknown events, adaptive techniques can be useful in order to improve models and be able to react to changes or unknown patterns and maintain acceptable performance for classification algorithms. Adaptive techniques are also important for online data analysis or streaming processing, where an incremental processing approach can be applied to deal with high throughput data analysis.

Compressed Sensing (CS) aims to obtain the most relevant information of a dataset, what makes it useful for compression, signal reconstruction, image processing and classification, such as in sparse representation classification (SRC) problems. Dictionary learning is related to CS but seeks to obtain compact and meaningful signal representations, known as dictionary, from learning algorithms or models. DL is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries. DL is based on the principle that some observations can be described by a sparse subset of atoms taken from a redundant dictionary, which represents the causes of some observations of the world. The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, image processing, data decomposition, feature extraction and classification \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.

Dictionaries are either available analytically, or can be learned from a suitable training set. DL can be formulated as an adaptive algorithm that learns from training data that is updated to represent changes or to adapt to new patterns. Iterative dictionary learning has been used as a blind technique for feature extraction, improving classification without require feature selection or principal component analysis

@@

Considering that big data problems require techniques to deal with multidimensional data in order to make sense of structure and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highlight discriminative structure and relationships of one dataset, we also propose a tensor-based dictionary learning method for fraud detection from imbalanced data, in order to apply the tensor decomposition for dictionary learning methods to evidentiate the discriminative sensing of a fraud detection dataset. Specifically, we propose to apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification of mobile money transactions. We propose to use tensor-based dictionary learning for learning fraudulent and legitimate data separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by some metrics.

@@

We analyze problems related to detection of information security issues and propose new approaches to improve malicious behavior detection through signal processing techniques. The results of the work presented in this thesis provide the following contributions:

\begin{enumerate}
	\item We propose an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 dataset;
	\item We discuss the computational complexity of the proposed framework and evaluate the required processing time for tested scenarios;
	\item We propose an architecture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
	\item We discuss the processing time of the proposed framework for mobile devices;
	\item We propose a tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
\end{enumerate}

@@

This thesis is organized as follows. In Chapter \ref{ch:2_mos_eig_sim}, we propose a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks, based on eigenvalue analysis, model order selection (MOS) and similarity analysis. Chapter \ref{ch:3_mobile} presents an evaluation of an approach and architecture based on user behavior analysis through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats in a mobile application. Chapter \ref{ch:4_tensor_dl} proposes a tensor-based dictionary learning approach for fraud detection from mobile payment transactions. Chapter \ref{ch:5_conclusionfuturework} draws the conclusions and the suggestions for future work. Furthermore, the Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes and the Appendix \ref{apx:b_csf_fs} presents a critical factors analysis based on Principal Component Analysis (PCA) for visual discriminant analysis, and presents an approach based on Recursive Feature Elimination (RFE) combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of Brazilian public organizations, in order to identify the Critical Success Factors (CSF) for IT governance of the public sector according to TCU.

### 2018-05-15 docs/thesis/chapters/2_eigen_similarity/eigen_similarity.tex
new project for adaptive multivariage gaussian applied to network attack detection

According to flood and port scan attacks' behavior, flood attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively. These characteristics are substantiated by the results obtained through the analysis based on sample covariation of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:2_experimentalresults}, which shows that the main components of flood attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.

@@

Therefore, $\boldsymbol{v}^{(q)}$ should be calculated from the last $\boldsymbol{X}^{(q)}$ time frame without attack, and $\boldsymbol{v}_{(m,\hat{n})}$ should be calculated from the same traffic appened of all $n$-th minutes until the identified minute under attack, denoted as $\boldsymbol{X}_n$.

@@

Table \ref{tab:tab12} presents the evaluated results for attack detection, considering rates of TP, FP \cite{fleiss2013statistical} and misclassification, which is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$ \cite{bhuyan2014network}. The analysis based on sample covariation of zero mean variables is evaluated for flooding behavior of netpune attacks, obtaining rates of 100.00 \% for true positive (TP) detection and 60.00 \% for false positive (FP) detection from 30 time frames. The results also show 50.00 \% of misclassification rate , which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed. 
The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA dataset presents high number of packets per time  from one source to one target, with no variation on IP source or target port. This observation corroborates with previous evaluations of the DARPA dataset that highlight issues regarding traffic redundancy.

>>>>

According to flood and port scan attacks' behavior, flood attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively. These characteristics are substantiated by the results obtained through the analysis based on sample covariance of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:2_experimentalresults}, which shows that the main components of flood attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.

@@

Therefore, $\boldsymbol{v}^{(q)}$ should be calculated from the last $\boldsymbol{X}^{(q)}$ time frame without attack, and $\boldsymbol{v}_{(m,\hat{n})}$ should be calculated from the same traffic appended of all $n$-th minutes until the identified minute under attack, denoted as $\boldsymbol{X}_n$.

@@

Table \ref{tab:tab12} presents the evaluated results for attack detection, considering rates of TP, FP \cite{fleiss2013statistical} and misclassification, which is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$ \cite{bhuyan2014network}. The analysis based on sample covariance of zero mean variables is evaluated for flooding behavior of netpune attacks, obtaining rates of 100.00 \% for true positive (TP) detection and 60.00 \% for false positive (FP) detection from 30 time frames. The results also show 50.00 \% of misclassification rate, which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed. The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA dataset presents high number of packets per time from one source to one target, with no variation on IP source or target port. This observation corroborates with previous evaluations of the DARPA dataset that highlight issues regarding traffic redundancy.

### 2018-05-15 docs/thesis/chapters/5_conclusion_future_work/conclusion.tex
new project for adaptive multivariage gaussian applied to network attack detection

The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis. The main contributions of this thesis were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.

@@

This thesis shows that a tensor-based approach can be used for fraud detection from mobile money transactions, with average results when compared to well known classifiers in one scenario with highly skewed datasets.

@@

\begin{enumerate}
	\item We proposed an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 dataset;
	\item We discussed the computational complexity of the proposed framework and evaluated the required processing time for tested scenarios;
	\item We proposed an architecture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
	\item We discussed the processing time of the proposed framework for mobile devices;
	\item We proposed a tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
	\item We published the following papers reporting our results:
	\begin{enumerate}
		\item T. P. B. Vieira, D. F. Ten\'orio, J. P. C. L. da Costa, E. P. de Freitas, G. Del Galdo, and R. T. de Sousa J\'unior, \textit{Model Order Selection and Eigen Similarity based Framework for Detection and Identication of Network Attacks}. Journal of Networking and Computer Applications (JNCA), Vol 90, Jul 2017, Pages 26–41 \cite{vieira2017model}.
		\item T. Galibus, T. P. B. Vieira, E. P. de Freitas, R. Albuquerque, J. P. C. L. da Costa, R. T. de Sousa Jr, V. Krasnoproshin, A. Zaleski, H. Vissia, and G. del Galdo, \textit{Offline Mode for Corporate Mobile Client Security Architecture}. Mobile Networks and Applications, Springer, Mar 2017, Pages 1-17 \cite{galibus2017offline}.
		\item K. H. C. Ramos, R. T. Sousa Jr, T. P. B. Vieira, and J. P. C. L. da Costa, \textit{Discovering Critical Success Factors for Information Technologies Governance through Bibliometric Analysis of Research Publications in This Domain}. Information (Yamaguchi), 2016 \cite{ramos2016information}.
		\item K. H. C. Ramos, T. P. B. Vieira, J. P. C. L. da Costa, and R. T. Sousa Jr., \textit{ANÁLISE MULTIDIMENSIONAL DE FATORES CRÍTICOS DE SUCESSO EM GOVERNANÇA DE TI NA ADMINISTRAÇÃO PÚBLICA FEDERAL À LUZ DOS DADOS DE CONTROLE EXTERNO}. in Proc. 12th International Conference on Management of Technology and Information Systems (CONTECSI), 2015, São Paulo, Brazil \cite{}.
	\end{enumerate}
\end{enumerate}

>>>>

The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis. The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.

@@

Finally, this thesis shows that a tensor-based approach can be used for fraud detection from mobile money transactions, with average results when compared to well known classifiers in one scenario with highly skewed datasets.

@@

\begin{enumerate}
	\item We proposed an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 dataset;
	\item We discussed the computational complexity of the proposed framework and evaluated the required processing time for tested scenarios;
	\item We proposed an architecture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
	\item We discussed the processing time of the proposed framework for mobile devices;
	\item We proposed a tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
	\item We published the following papers reporting our results:
	\begin{enumerate}
		\item T. P. B. Vieira, D. F. Ten\'orio, J. P. C. L. da Costa, E. P. de Freitas, G. Del Galdo, and R. T. de Sousa J\'unior, \textit{Model Order Selection and Eigen Similarity based Framework for Detection and Identication of Network Attacks}. Journal of Networking and Computer Applications (JNCA), Vol 90, Jul 2017, Pages 26–41 \cite{vieira2017model}.
		\item T. Galibus, T. P. B. Vieira, E. P. de Freitas, R. Albuquerque, J. P. C. L. da Costa, R. T. de Sousa Jr, V. Krasnoproshin, A. Zaleski, H. Vissia, and G. del Galdo, \textit{Offline Mode for Corporate Mobile Client Security Architecture}. Mobile Networks and Applications, Springer, Mar 2017, Pages 1-17 \cite{galibus2017offline}.
		\item K. H. C. Ramos, R. T. Sousa Jr, T. P. B. Vieira, and J. P. C. L. da Costa, \textit{Discovering Critical Success Factors for Information Technologies Governance through Bibliometric Analysis of Research Publications in This Domain}. Information (Yamaguchi), 2016 \cite{ramos2016information}.
		\item K. H. C. Ramos, T. P. B. Vieira, J. P. C. L. da Costa, and R. T. Sousa Jr., \textit{Análise Multidimensional de Fatores Críticos de Sucesso em Governança de TI na Administração Pública Federal à Luz dos Dados de Controle Externo}. in Proc. 12th International Conference on Management of Technology and Information Systems (CONTECSI), 2015, São Paulo, Brazil \cite{ramos2015}.
	\end{enumerate}
\end{enumerate}
